---
title: "Inspect profiles"
subtitle: "cmQTL project"
author: "Shantanu Singh"
date: "July 2020"
output: github_document
---

```{r echo=FALSE}
show_table <- print
```

If running interactively in RStudio, 

- change `output` in the header of this markdown to `html_notebook` and
- change to `eval=TRUE` below

When knitting for pushing to GitHub,

- change `output` in the header of this markdown to `github_document` and
- change to `eval=TRUE` below

```{r eval=TRUE}
show_table <- knitr::kable
```

```{r}
library(ggplot2)
library(glue)
library(magrittr)
library(tidyverse)  
```

# Read profiles

```{r}
plates <- c("BR00106708", 
            "BR00106709", 
            "BR00107338", 
            "BR00107339", 
            "cmqtlpl1.5-31-2019-mt", 
            "cmqtlpl261-2019-mt",
            "cmQTLplate7-7-22-20"
)

plate_id <- plates[[1]]

profiles <- 
  plates %>%
  map_df(~read_csv(file.path("profiles", paste0(.x, "_normalized.csv")), 
                   col_types = cols(
                     .default = col_double(),
                     Metadata_Plate = col_character(),
                     Metadata_Well = col_character(),
                     Metadata_Assay_Plate_Barcode = col_character(),
                     Metadata_Plate_Map_Name = col_character(),
                     Metadata_well_position = col_character()
                   ),
                   progress = FALSE))

```

`Metadata_line_ID` are empty wells or those that failed

```{r}
profiles %<>%
  filter(Metadata_line_ID != 0)
```

# Inspect profiles

Across how many plates is each cell line plated?

```{r}
profiles %>%
  distinct(Metadata_Plate, Metadata_line_ID) %>%
  group_by(Metadata_line_ID) %>%
  tally(name = "n_plates") %>%
  group_by(n_plates) %>%
  tally(name = "n_lines") %>%
  show_table
```

List cell lines that are present on more than one plate

```{r}
profiles %>%
  distinct(Metadata_Plate, Metadata_line_ID) %>%
  group_by(Metadata_line_ID) %>%
  tally(name = "n_plates") %>%
  filter(n_plates == 2) %>%
  inner_join(profiles %>% distinct(Metadata_Plate, Metadata_line_ID)) %>%
  ungroup() %>%
  arrange(Metadata_line_ID) %>%
  show_table
```

