---
title: "Inspect cmQTLplate7-2-27-20"
output: github_document
---

```{r echo=FALSE}
show_table <- print
```

If running interactively in RStudio, 

- change `output` in the header of this markdown to `html_notebook` and
- change to `eval=TRUE` below

When knitting for pushing to GitHub,

- change `output` in the header of this markdown to `github_document` and
- change to `eval=FALSE` below

```{r eval=TRUE}
show_table <- knitr::kable
```


```{r}
suppressPackageStartupMessages(library(magrittr))
suppressPackageStartupMessages(library(tidyverse))
```


This chunk was run on AWS 

Randomly sample ~5000 cells

```{r eval=FALSE}
sqlite_file <- "../../../backend/2020_03_05_Batch6/cmQTLplate7-2-27-20/cmQTLplate7-2-27-20.sqlite"

db <- DBI::dbConnect(RSQLite::SQLite(), sqlite_file, loadable.extensions = TRUE)

nuclei <- 
  DBI::dbGetQuery(db, "SELECT * FROM Nuclei ORDER BY RANDOM() LIMIT 5000;") %>%
  collect() %>%
  select(-ObjectNumber)

nuclei_lut <-
  nuclei %>%
  select(TableNumber, ImageNumber, Nuclei_Number_Object_Number)

cells <- 
  tbl(src = db, "Cells") %>% 
  select(-ObjectNumber) %>% 
  inner_join(
    nuclei_lut,
    by = c("TableNumber", 
           "ImageNumber", 
           "Cells_Parent_Nuclei" = "Nuclei_Number_Object_Number"),
    copy = TRUE
  ) %>%
  collect()

cytoplasm <- 
  tbl(src = db, "Cytoplasm") %>% 
  select(-ObjectNumber) %>% 
  inner_join(
    nuclei_lut,
    by = c("TableNumber", 
           "ImageNumber", 
           "Cytoplasm_Parent_Nuclei" = "Nuclei_Number_Object_Number"),
    copy = TRUE
  ) %>%
  collect()

nuclei_cells <- 
  inner_join(
    nuclei,
    cells,
    by = c("TableNumber", 
           "ImageNumber", 
           "Nuclei_Number_Object_Number" = "Cells_Parent_Nuclei")
  )

nuclei_cells_cytoplasm <- 
  inner_join(
    nuclei_cells,
    cytoplasm,
    by = c("TableNumber",
           "ImageNumber",
           "Cells_Number_Object_Number" = "Cytoplasm_Parent_Cells")
  )

nuclei_cells_cytoplasm %>% write_csv("cmQTLplate7-2-27-20_sampled.csv.gz")
```

Load the subsample

```{r}
sampled_cells <- read_csv("data/cmQTLplate7-2-27-20_sampled.csv.gz")
```

Specify which features are going to be check for NA
`Area` features are excluded because we are going to use them to check whether they may be able to predict whether a feature is `NA`

```{r}
no_check_na_features <- 
  c("TableNumber", "ImageNumber", 
      "Cells_AreaShape_Area",
      "Cytoplasm_AreaShape_Area",
      "Nuclei_AreaShape_Area")

check_na_features <- 
  setdiff(
    names(sampled_cells),
    no_check_na_features
  )
```

Summarize number of cells with NA values per feature

```{r}
na_frequency <- 
  sampled_cells %>% 
  summarize_at(check_na_features, ~sum(is.na(.))) %>%
  pivot_longer(everything(), values_to = "number_of_na")
```


```{r}
na_frequency %>%
  arrange(desc(number_of_na)) %>%
  filter(number_of_na >= 0) %>%
  write_csv("data/plate7_na_features.csv")
```


```{r}
na_frequency %>%
  filter(number_of_na >= 10) %>%
  arrange(desc(number_of_na)) %>%
  show_table
```

Report whether a feature value is NA per cell

```{r}
na_flag <- 
  sampled_cells %>% 
  mutate_at(check_na_features, ~as.double(is.na(.)))
```

For features that have more than 10 cells as `NA`, how many are non-correlation features?
(none)

```{r}
na_flag %>% 
  summarize_at(check_na_features, sum) %>%
  pivot_longer(everything(), values_to = "number_of_na") %>%
  filter(number_of_na >= 10) %>%
  arrange(desc(number_of_na)) %>% 
  filter(!str_detect(name, "_Correlation_")) %>%
  show_table
```

Pick the feature with the most number of `NA`s and check whether it can be explained by Area features
(nope)

```{r}
data_matrix <- 
  na_flag %>%
  mutate(Nuclei_Correlation_Costes_AGP_Mito = scale(Nuclei_Correlation_Costes_AGP_Mito)) %>%
  select(one_of(c(no_check_na_features, "Nuclei_Correlation_Costes_AGP_Mito"))) %>%
  select(-TableNumber, -ImageNumber)

model <- lm(Nuclei_Correlation_Costes_AGP_Mito ~ ., data_matrix)

summary(model)
```

Check other features

```{r}
data_matrix <- 
  na_flag %>%
  mutate(Cells_Correlation_Costes_AGP_Mito = scale(Cells_Correlation_Costes_AGP_Mito)) %>%
  select(one_of(c(no_check_na_features, "Cells_Correlation_Costes_AGP_Mito"))) %>%
  select(-TableNumber, -ImageNumber)

model <- lm(Cells_Correlation_Costes_AGP_Mito ~ ., data_matrix)

summary(model)
```

For what 

```{r}
sorted_features_by_na <- 
  na_frequency %>%
  filter(number_of_na >= 10) %>%
  arrange(desc(number_of_na))
```


```{r}
na_flag_long <- 
  na_flag %>%
  select(-one_of(no_check_na_features)) %>%
  select(-matches("_Costes_|_RWC_|_Manders_|_K_")) %>%
  rownames_to_column("cell_id") %>%
  mutate(cell_id = as.integer(cell_id)) %>%
  pivot_longer(-cell_id)
```


```{r}
na_counts_per_cell <-
  na_flag_long %>%
  group_by(cell_id) %>%
  summarize(value = sum(value)) %>%
  arrange(desc(value))
```


```{r}
na_counts_per_cell %>%
  slice(1:20) %>%
  show_table
```


```{r}
na_counts_per_cell %>% 
  filter(value < 100) %>%
  ggplot(aes(value)) + geom_histogram(binwidth = 1)
```


```{r}
na_counts_per_cell %>% 
  filter(between(value, 1, 100)) %>%
  ggplot(aes(value)) + geom_histogram(binwidth = 1)
```

```{r}

bad_cells <- 
  na_counts_per_cell %>% 
  filter(value > 100)

na_flag_long %>% 
  anti_join(bad_cells) %>%
  filter(value == 1) %>%
  select(name) %>%
  group_by(name) %>%
  tally() %>%
  arrange(desc(n))
```

```{r}
cells_with_na <- 
  na_flag_long %>% 
  anti_join(
    na_counts_per_cell %>% 
      filter(value > 100) %>%
      select(cell_id)
  ) %>%
  group_by(cell_id) %>%
  summarize(has_na_feature = sum(value) > 0) %>%
  filter(has_na_feature)
```

First remove these many "bad" cells (many features are `NA`)

```{r}
bad_cells %>% count %>% show_table
```

Next remove all Correlation features, then report how many cells have any feature as `NA`

```{r}
cells_with_na %>% count %>% show_table
```

